Impact of grids and dynamical cores in CESM2.2 on the surface mass balance of the Greenland Ice Sheet

### 2022-09-08 first_review/response-to-reviewers.tex

\section*{Reviewer 1}

\subsection*{General comments}

Given that the different grid and dycore choices come with considerable different computational costs, I think it would be useful to include a section addressing that specific question. Some information touching on the costs appears at different places in the manuscript. It would be good to bring that all together somewhere and make a clear comparison, e.g. including a comparative table. \newline

{\color{blue}{We have added a subsection on computational costs to address the reviewers concern, in the methods section.}} \newline

@@

The way the term surface mass balance (SMB) is defined and used in the manuscript is confusing. The different components to the SMB are not consistently applied and there are two different definitions of the SMB in use. In part this confusion is related to the fact that the model itself does not calculate all the SMB components and assumptions have to be made about the separation between snow and ice, especially in the currently discussed uncoupled case. I would suggest to start with a physical definition of SMB with all its components in an equation and then explain based on that what the model is actually doing and what constraints have to be applied. In this context it may help to describe what the model would do in a coupled setup with active CISM and then apply further constraints for the diagnostic case currently described. \newline

{\color{blue}{We thanks the reviewer for bringing this to our attention, and agree that our constant referencing of what the SMB refers to has led to some inconsistent definitions and confusion in the text. We have (1) removed attempts (other than very generally) to define the SMB (and the ablation zone) in parenthesis after introducing these terms in the introduction, (2) clarified the description of the SMB that is computed in CESM (the section ``Simulated surface mass balance"), and (3) made clear in the ``SMB analysis" section that the SMB equation we are computing in this paper is different from that computed by CESM.}} \newline

@@

In the discussion about the different dycores, I was wondering if there could be a notion that the latitude-longitude approaches are more mature because they have a longer history. Maybe further improvements on technical aspects like filtering and optimal solutions may be expected for the FE approaches. \newline

{\color{blue}{This is a good point. A case in point would be the improvements to the flow over topography going from CESM2.0 to CESM2.2. We are still developing CAM-SE. We have elaborated on this point in the methods section (where we point to the CESM2.2 enhancements in the Appendix) and the conclusions section, the second to last paragraph.}} \newline

@@

The relationship between atmospheric grid and land grid seems to be important. Would it be possible to say that higher resolution of the atmospheric grid improves the precipitation, while increased resolution of the land model improves the melt calculation/ablation components? \newline

{\color{blue}{Based on my understanding of how the atmosphere and land models are coupled, I don't think running the land model at a higher resolution (and holding the atmosphere resolution fixed) will alleviate the melt biases that tends to occur with partially ice covered grid cells. This is because it is the averaging of the land state over the atmospheric grid cell that is responsible for homogenizing the state. Even if the land cells are a lot smaller, and even binary in ice cover, If the fractional ice cover within an \textit{atmospheric grid cell} is still small, then we think that the atmospheric state that is then passed back to the land model will still be unduly influenced by the non-ice portions of the atmosphere grid cell.}}

@@

l174 Is it correct to say the dycore is a model? \newline

{\color{blue}{Reworded.}} \newline

l203 "the SE numerics have been enhanced". Enhanced compared to what. An earlier model version, I suspect. Mention that. \newline

{\color{blue}{Added: relative to the CESM2.0 release.}} \newline

@@

l525 "underline"? \newline

{\color{blue}{Yes. Fixed.}} \newline

l528 Did you test how much biased the results could be due to the relatively short period of 20 years in view of the large inter-annual variability? If you don't have more years, you could remove any of the 20 years and see how much the numbers change. \newline

{\color{blue}{We've done performed a Monte Carlo analysis similar to what is suggested and have found that the errors due to inter-annual variability (in Figure 10) are much smaller than our current error bars, and have therefore omitted these estimates.}} \newline

@@

l1071 Refer to final paper. https://doi.org/10.5194/tc-13-1547-2019 \newline

{\color{blue}{Done.}}

\section*{Reviewer 2}

The manuscript is well-written and provides an important contribution to understanding simulation of ice sheet SMB in global climate model simulations. The paper provides important documentation of the impact of changes to the grid and spatial resolution in CESM on ice sheet SMB. Some points below should be addressed before the manuscript is published but I feel these are minor overall. \newline

One general comment: surface albedo has been shown to be an important factor in climate model simulations of GrIS SMB (e.g. van Angelen et al., 2012; Cullather et al., 2014; Helsen et al., 2017; Alexander et al., 2019), but this has not been discussed by the authors. Some of the melt differences could be influenced by the interaction between precipitation, surface albedo, and melt. In areas of underestimated snowfall, which tend to occur along the ice sheet margins, erroneous bare ice exposure can lead to a darker surface and enhanced melt. IThe authors should examine the possibility of differences in albedo influencing simulated melt. \newline

%Cullather, R. I., Nowicki, S. M. J., Zhao, B., & Suarez, M. J. (2014) Evaluation of the surface representation of the Greenland Ice Sheet in a General Circulation Model. J. Climate, 27 (13), 4835-4856. \newline

%Helsen, M. M., van de Wal, R. S. W., Reerink, T. S., Bitanja, R., Madsen, M. S., Yang, S., Li, Q., & Zhang, Q. (2017) On the importance of the albedo parameterization for the mass balance of the Greenland ice sheet in EC-Earth. The Cryosphere, 11, 1949-2017. \newline

%Alexander, P. M., LeGrande, A. N., Fischer, E., Tedesco, M., Fettweis, X., Kelley, M., Nowicki, S. M. J., & Schmidt, G. A. (2019) Simulated Greenland surface mass balance in the GISS ModelE2 GCM: Role of the ice sheet surface. J. Geophys. Res., 124 (3), 750- 765.

1. Lines 19-24: The language is confusing here. The quasi-uniform grids should be introduced with a brief description of how they are constructed, and what “quasi- uniform” means. Then their performance in terms of surface mass balance can be discussed, followed by the impact of grid refinement on the simulation. Also perhaps it can be mentioned here how one dycore is linked to two lon-lat configurations and another is linked to the quasi-uniform unstructured grid. \newline

2. Line 38: Again, please explain the meaning of “quasi-uniform unstructured grid”  \newline

3. Lines 52-55: Again, it would be helpful if “globally uniform unstructured grid”, “dynamical cores”, and “variable-resolution grids” can be briefly explained. Additionally, perhaps it should be explained here that two dycores are examined, one linked to the longitude latitude grid, and the other linked to unstructured grids.  \newline

4. Line 94: Why is “conversely” used here? It seems like a word such as “Additionally” would be more appropriate.  \newline

5. Lines 115-116: Clarify that this is a positive precipitation bias in the interior.  \newline

6. Lines 111-123: Here there is no mention of elevation classes. I suggest mentioning them briefly here as this is a way to increase SMB resolution (from the standpoint of runoff but not precipitation) that is used in CESM  \newline
.
7. Figure 1, caption: Please provide a brief description of each of the grids here for the benefit of the reader. E.g. (a) 2 latitude-longitude grid (f19), etc.  \newline

8. Line 189: Briefly explain “cubed-sphere”.  \newline

9. Table 1: It would be helpful to specify which grids are latitude-longitude and which are unstructured either as a column in the table or text in the caption.  \newline

10. Lines 229-230: Does ne30 indicate 30 quadrilateral elements per face and pg3 indicate 3 control volumes per element? Please clarify. \newline

11. Lines 239-243: I suppose degree coordinates here correspond the latitude dimension, or degrees at the equator. Please clarify.  \newline

12. Line 249: This is the first time RACMO is mentioned. Please explain briefly what RACMO is here.  \newline

13. Line 254: Clarify that “expensive” means “computationally expensive.” \newline

14. Line 301: Though not essential it would be interesting to know the lapse rate that is assumed. \newline

15. Line 314: I’m confused as to why accumulation below 10 m does not contribute to SMB. How is this addition of mass accounted for? \newline

16. Lines 318-319: What is the period for the prescribed SSTs? \newline

17. Lines 322-338: This section only includes discussion of RACMO and in-situ measurements, which are not used, while ERA5, CERES and CALIPSO are not at all discussed. These validation products must be documented here as well. \newline

18. Line 322: RACMO2 is still a model, and does not fall under an “observational” dataset. I suggest changing this to read “Validation Datasets”, “Evaluation Datasets” or “Comparison Datasets”. \newline

19. Figure 3 caption: I believe the left and right figures are switched here. Also, the fact that RACMO RCM simulations are included should be mentioned. \newline

20. Formula 1: The f’s in runoff seem to have been turned into function symbols. Also I suggest using “evaporation + sublimation” to avoid confusion with division. \newline

21. Line 406: I believe this should read “expressed as temperature differences”. 22. Lines 408-416, 429-433: How can the effect be primarily ascribed to condensational heating if the temperature differences do not align in magnitude or spatial extent? Are there feedbacks involved? Please clarify in the text, or qualify the statements. \newline

23. Lines 437-438: This also seems to be true in other cases. \newline

24. Lines 440-441: This seems consistent with other results that increasing resolution led to a temperature increase. \newline

25. Line 459: Perhaps the authors could mention the potential influence of clouds earlier as prior to this point, they are not mentioned. \newline

26. Line 479: Clarify that cooler temperatures in the coarse resolution simulations are being referred to here. \newline

27. Figure 8: The CALIPSO grid appears to be coarser resolution than the CERES grid. Is this a 2 grid? \newline

28. Lines 573-575: Wouldn’t the elevation class scheme fix this problem by raising the ice-covered sub-grid portion to a higher elevation? \newline

29. Lines 606-616: If the biases are larger at low elevations, it could also be coincidental that these areas also have lower ice fractions. I think the authors should also consider the possibility of other factors. \newline

30. Lines 675-680: Perhaps the fact that clouds and precipitation move more easily inland could reduce snowfall and increase shortwave radiation along the coast, enhancing melt. \newline

31. Lines 681-688: Again, I’m curious about the impact of elevation classes and whether precipitation-albedo feedbacks could influence the melt biases. I suppose the elevation classes might not entirely balance out other coarse- resolution effects. \newline

32. Line 719: Change “spectra-element” to “spectral-element”. \newline

33. Line 729: g can be defined after mentioning T1 for clarity. \newline

34. Line 731: Add “is the” before “surface geopotential”. \newline

35. Equation A5: How is T(ref) computed? \newline

36. Lines 740-748: Here some terms should be defined, including Tv, qv and p0.

### 2022-09-08 second_submit/arcticgrids.tex
working on reviewer 2

General Circulation Models (GCMs) are powerful tools for understanding the meteorology and climate of the high latitudes, which are among the most sensitive regions on Earth to global and environmental change. GCMs differ vastly in their numerical treatment of polar regions because of the so-called \textit{pole problem} \cite{W2007JMSJ}. The pole problem refers to numerical instability arising from the convergence of meridian lines into polar singularities on latitude-longitude grids (e.g., Figure~\ref{fig:uni-grids}a, hereafter referred to as \textit{lat-lon} grids). Depending on the numerics, methods exist to suppress this instability, and lat-lon grids may be advantageous for polar processes by representing structures with finer resolution than elsewhere in the computational domain. With the recent trend towards quasi-uniform unstructured grids, any potential benefits of lat-lon grids in polar regions may be lost {\color{blue}{(hereafter, {\textit{quasi-uniform}} refers to approximately isotropic grids, which is not a characteristic of lat-lon grids, which are highly anisotropic)}}. In this study, we evaluate a number of grids and dynamical cores (hereafter referred to as \textit{dycores}) available in the Community Earth System Model, version 2.2 \cite<CESM2.2;>{CESM2}, including new variable-resolution grids, to understand their impacts on the simulated Arctic climate. We focus specifically on the climate and surface mass balance of the Greenland Ice Sheet.

In the 1970s, the pole problem was largely defeated through the adoption of efficient spectral transform methods in GCMs \cite<see>[and references therein]{W2007JMSJ}.
These methods transform grid point fields into a global, isotropic representation in wave space, where linear operators (e.g., horizontal derivatives) in the (truncated) equation set can be solved exactly. While spectral transform methods are still used today, local numerical methods have become desirable for their ability to run efficiently on massively parallel systems. The pole problem has thus re-emerged in contemporary climate models that use lat-lon grids, and some combination of reduced grids (modified lat-lon grids, with cells elongated in the longitudinal direction over the polar regions) and polar filters are necessary to ameliorate this numerical instability \cite{JW2010LNCSE}. Polar filters subdue the growth of unstable computational modes by applying additional damping to the numerical solution over polar regions. This damping reduces the effective resolution in polar regions such that the resolved scales are \textit{approximately} the same everywhere on the grid. We emphasize \textit{approximately}, since it is conceivable that marginal increases in effective resolution occur over polar regions in lat-lon grids, despite polar filtering, since resolved waves can be represented with more grid points than at lower latitudes.

@@

The atmospheric component of CESM2.2 \cite{CESM2}, the Community Atmosphere Model, version 6.3 \cite<CAM6;>{GETAL2019GRL,CAM63}, supports several different atmospheric dynamical cores. These include dycores on lat-lon grids, such as finite-volume \cite<FV;>{L2004MWR} and Eulerian spectral transform \cite<EUL;>{CETAL006JC} models, and dycores built on unstructured grids, including spectral-element \cite<SE;>{LetAl2018JAMES} and finite-volume 3 \cite<FV3;>{PL2007JCP} models. This study compares the performance of the SE and FV dycores, omitting the EUL and FV3 dycores. CESM2 runs submitted to CMIP6 used the FV dycore, whereas the SE dycore is often used for global high-resolution simulations \cite<e.g., >{SETAL2014JAMES,BETAL2018CC,IHESP} due to its higher throughput on massively parallel systems \cite{DetAl2012IJHPCA}.

@@

In CESM2.2, the SE numerics have been enhanced {\color{blue}{relative to the CESM2.0 release}} to mitigate spurious noise over topography. These algorithmic changes are described in \ref{sec:a1}.

The SE dycore supports regional grid refinement via its VR configuration, requiring two enhancements over {\color{blue}{quasi-uniform resolution grids}}. First, as the numerical viscosity increases with resolution, explicit hyperviscosity relaxes according to the local element size, reducing in strength  by an order of magnitude per halving of the grid spacing. A tensor-hyperviscosity formulation is used \cite{GetAl2014GMD}, which adjusts the coefficients in two orthogonal directions to more accurately target highly distorted quadrilateral elements. Second, the topography boundary conditions are smoothed in a way that does not excite grid scale modes, and so the NCAR topography software \cite{gmdd-8-4623-2015} has been modified to scale the smoothing radius by the local element size, resulting in rougher topography in the refinement zone.

@@

 \begin{table*}
 \centering
 \scriptsize
 \begin{tabular}{lcccccc}
   \hline
   grid name & dycore & $\Delta x_{\mathrm{eq}}$ (km) & $\Delta x_{\mathrm{fine}}$ (km) & $\Delta t_{\mathrm{phys}}$ (s) & {\color{blue}{25 nodes, cost}} & {\color{blue}{50 nodes, cost}} \\ 
   \hline
   $\tt f19$ & FV & 278 & - &1800 & {\color{blue}{436.66}} & {\color{blue}{-}} \\
   $\tt f09$ & FV & 139 & - &1800 & {\color{blue}{1534.57}} & {\color{blue}{2024.24}} \\
   $\tt ne30pg2$ & SE-CSLAM & 167 & - & 1800 & {\color{blue}{1497.26}} & {\color{blue}{1683.97}} \\
   $\tt ne30pg3$ & SE-CSLAM & 111 & - & 1800 & {\color{blue}{1890.48}} & {\color{blue}{2090.43}} \\
   $\tt ne30pg3^{*}$ & SE-CSLAM & 111 & - & 450 & {\color{blue}{-}} & {\color{blue}{-}} \\
   $\tt Arctic$ & SE & 111 & 28 & 450 & {\color{blue}{15947.41}} & {\color{blue}{16675.45}} \\
   $\tt Arctic-GrIS$ & SE & 111 & 14 & 225 & {\color{blue}{40305.03}} & {\color{blue}{41036.67}} \\
 \hline
 \end{tabular}
  \caption{Grids and dycores used in this study. $\Delta x_{\mathrm{eq}}$ is the average equatorial grid spacing, $\Delta x_{\mathrm{fine}}$ is the grid spacing in the refined region (if applicable), and $\Delta t_{\mathrm{phys}}$ is the physics time step. FV refers to the finite-volume dycore, SE the spectral-element dycore, and SE-CSLAM the spectral-element dycore with CSLAM tracer advection.  We use the $\tt ne30pg3$ grid for two runs with different values of $\Delta t_{\mathrm{phys}}$. {\color{blue}{The last columns provide the computational costs in core hours per simulated year (CHPSY). The costs are from single month runs using 25 nodes and 50 nodes on the Cheyenne supercomputer \cite{cheyenne}.}}}
 \label{tbl:table1}
 \end{table*}

@@

\begin{figure}[t]
\begin{center}
         \includegraphics[width=130mm]{temp_contours_diffCALIPSOdiffCERESdiffRACMO.pdf}
\end{center}
\caption{1979-1998 northern hemisphere summer surface incident shortwave radiation bias (W/m2), computed as the difference (top) from CERES, and (middle) RACMO2.3p2 dataset, and the (bottom) total cloud fraction bias relative to the CALIPSO dataset. CALIPSO and CERES differences are found by mapping the model output to the $1^{\circ}$ grid, and differences in the bottom panel are computed after mapping the RACMO2.3p2 dataset to the individual model grids. Note that the averaging period for the CALIPSO-GOCCP and CERES-EBAF panels, 2006-2017 and 2003-2020, respectively, are different from the averaging period for the model results.}
\label{fig:FSDS}
\end{figure}

\begin{figure}[t]
\begin{center}
         \includegraphics[width=130mm]{temp_contours_diffRACMO}
\end{center}
\caption{1979-1998 (top) annual precipitation and (bottom) ice/snow melt biases relative to RACMO2.3p2, evaluated on the native model grids. The precipitation biases are expressed as fractional changes, whereas the melt biases are absolute changes (mm/yr). In the bottom panel, the \citeA{RM2012GRL} basin boundaries are shown in grey for each model grid. Note that Figure~\ref{fig:basin} uses the basin boundaries for the two common ice masks, shown in the $\tt f19$ and $\tt ne30pg2$ panels, in computing the basin-scale integrals. Blue lines in the $\tt f19$ panel show the location of the two transects plotted in Figure~\ref{fig:ztrans}..}
\label{fig:prect}
\end{figure}

@@

The bottom panel of Figure~\ref{fig:prect} presents the biases in the combined ice/snow melt as map plots. These plots show that the largest melt biases are on the southeast and northwest coasts, where large coarse-grid cells overlap with the ocean. One possibility is that these problematic grid cells are situated at lower elevations than the true ice sheet surface, leading to a warm bias and too much melt. Figure~\ref{fig:ztrans} shows the representation of the ice sheet surface along two transects on the different grids, compared to the high-resolution dataset used to generate CAM topographic boundary conditions \cite{GMTED2010,gmdd-8-4623-2015}. The two transects are shown in Figure~\ref{fig:prect}: the east-west ``K-transect'' in southwest Greenland and a transect extending from the central dome down to the  Kangerlussuaq glacier on the southeast coast. The $1^{\circ}-2^{\circ}$ grids are noticeably coarse, with only a handful of grid cells populating the transect. The $\tt f09$ grid is a bit of an exception since the grid cells narrow in the zonal direction at high latitudes, and so a larger number of grid cells fit into the east-west transects. The VR grids more accurately reproduce the steep margins of the ice sheet, capturing the characteristic parabolic shape of the GrIS margin.

@@

Figure~\ref{fig:bias} shows the mean melt bias, relative to both RACMO datasets, conditionally sampled based on grid cell ice fraction in the GrIS region. Errors are computed after mapping the melt rates to the common ice masks using different methods, described in section~\ref{sec:SMB}. The grid cell ice fractions therefore pertain to ice fractions on the low-resolution common ice masks. Also shown are the $\pm 1$~standard deviation of the biases for each bin. The figure shows that coarser grids can be characterized by a monotonic increase in melt bias as the grid cell ice fraction decreases. The VR grids have the smallest melt biases for small grid cell ice fractions (smaller than 50\%), the {\color{blue}{quasi-}}uniform SE grids and $\tt f19$ have the largest melt biases and the $\tt f09$ grid melt biases lie between these two cases. Figure~\ref{fig:bias} generally supports our hypothesis that the prevalence of mixed-grid cells in the ablation zone of coarse grids is responsible for their large melt bias.

@@

@@

As modeling systems move away from lat-lon grids towards quasi-uniform unstructured grids, it is worth taking stock of whether this will degrade the simulated polar climate. We have found that the $1^{\circ}$ FV model has clear advantages over the $1^{\circ}$ SE model for simulating the GrIS SMB. That is, the simulated GrIS SMB will be adversely impacted in future CESM versions, after the FV dycore is phased out. This finding will not interrupt the ongoing transition towards unstructured grids in CESM, which is largely driven by gains in computational efficiency and grid refinement capabilities. We therefore provide the Arctic refined-meshes to the community by way of CESM2.2, providing users the option to simulate a more realistic GrIS SMB, although at a substantial computational premium relative to conventional $1^{\circ}-2^{\circ}$ grids.

>>>>

General Circulation Models (GCMs) are powerful tools for understanding the meteorology and climate of the high latitudes, which are among the most sensitive regions on Earth to global and environmental change. GCMs differ vastly in their numerical treatment of polar regions because of the so-called \textit{pole problem} \cite{W2007JMSJ}. The pole problem refers to numerical instability arising from the convergence of meridian lines into polar singularities on latitude-longitude grids (e.g., Figure~\ref{fig:uni-grids}a, hereafter referred to as \textit{lat-lon} grids). Depending on the numerics, methods exist to suppress this instability, and lat-lon grids may be advantageous for polar processes by representing structures with finer resolution than elsewhere in the computational domain. With the recent trend towards quasi-uniform unstructured grids, any potential benefits of lat-lon grids in polar regions may be lost {\color{blue}{(hereafter, {\textit{quasi-uniform}} refers to approximately isotropic grids, in contrast to lat-lon grids, which are highly anisotropic)}}. In this study, we evaluate a number of grids and dynamical cores (hereafter referred to as \textit{dycores}) available in the Community Earth System Model, version 2.2 \cite<CESM2.2;>{CESM2}, including new variable-resolution grids, to understand their impacts on the simulated Arctic climate. We focus specifically on the climate and surface mass balance of the Greenland Ice Sheet.

In the 1970s, the pole problem was largely defeated through the adoption of efficient spectral transform methods in GCMs \cite<see>[and references therein]{W2007JMSJ}.
These methods transform grid point fields into a global, isotropic representation in wave space, where linear operators (e.g., horizontal derivatives) in the (truncated) equation set can be solved exactly. While spectral transform methods are still used today, local numerical methods have become desirable for their ability to run efficiently on massively parallel systems. The pole problem has thus re-emerged in contemporary climate models that use lat-lon grids, and some combination of reduced grids (modified lat-lon grids, with cells near the polar singularity elongated in the zonal direction) and polar filters are necessary to ameliorate this numerical instability \cite{JW2010LNCSE}. Polar filters subdue the growth of unstable computational modes by applying additional damping to the numerical solution over polar regions. This damping reduces the effective resolution in polar regions such that the resolved scales are \textit{approximately} the same everywhere on the grid. We emphasize \textit{approximately}, since it is conceivable that marginal increases in effective resolution occur over polar regions in lat-lon grids, despite polar filtering, since resolved waves can be represented with more grid points than at lower latitudes.

@@

The atmospheric component of CESM2.2 \cite{CESM2}, the Community Atmosphere Model, version 6.3 \cite<CAM6;>{GETAL2019GRL,CAM63}, supports several different atmospheric dynamical cores. These include dycores on lat-lon grids, such as finite-volume \cite<FV;>{L2004MWR} and Eulerian spectral transform \cite<EUL;>{CETAL006JC} models, and dycores built on unstructured grids, including spectral-element \cite<SE;>{LetAl2018JAMES} and finite-volume 3 \cite<FV3;>{PL2007JCP} models. This study compares the performance of the SE and FV dycores, omitting the EUL and FV3 dycores. CESM2 runs submitted to CMIP6 used the FV dycore, whereas the SE dycore is often used for global high-resolution simulations \cite<e.g., >{SETAL2014JAMES,BetAl2016CC,IHESP} due to its higher throughput on massively parallel systems \cite{DetAl2012IJHPCA}.

@@

{\color{blue}{SE is a next generation of dycore, and is less mature than the FV dycore due to its shorter history.}} In CESM2.2, the SE numerics have been enhanced {\color{blue}{relative to CESM2.0}} to mitigate spurious noise over topography. These algorithmic changes are described in \ref{sec:a1}. {\color{blue}{Future versions of CESM will likely continue to include further optimizations and enhancements to the (nascent) SE dycore.}}

The SE dycore supports regional grid refinement via its VR configuration, requiring two enhancements over {\color{blue}{quasi-uniform resolution grids}}. First, as the numerical viscosity increases with resolution, explicit hyperviscosity relaxes according to the local element size, reducing in strength  by an order of magnitude per halving of the grid spacing. A tensor-hyperviscosity formulation is used \cite{GetAl2014GMD}, which adjusts the coefficients in two orthogonal directions to more accurately target highly distorted quadrilateral elements. Second, the topography boundary conditions are smoothed in a way that does not excite grid scale modes, and so the NCAR topography software \cite{LetAl2015GMD} has been modified to scale the smoothing radius by the local element size, resulting in rougher topography in the refinement zone.

@@

 \begin{table*}
 \centering
 \scriptsize
 \begin{tabular}{lcccccc}
   \hline
   grid name & dycore & $\Delta x_{\mathrm{eq}}$ (km) & $\Delta x_{\mathrm{fine}}$ (km) & $\Delta t_{\mathrm{phys}}$ (s) & {\color{blue}{cost(25 nodes)}} & {\color{blue}{cost(50 nodes)}} \\ 
   \hline
   $\tt f19$ & FV & 278 & - &1800 & {\color{blue}{436.66}} & {\color{blue}{-}} \\
   $\tt f09$ & FV & 139 & - &1800 & {\color{blue}{1534.57}} & {\color{blue}{2024.24}} \\
   $\tt ne30pg2$ & SE-CSLAM & 167 & - & 1800 & {\color{blue}{1497.26}} & {\color{blue}{1683.97}} \\
   $\tt ne30pg3$ & SE-CSLAM & 111 & - & 1800 & {\color{blue}{1890.48}} & {\color{blue}{2090.43}} \\
   $\tt ne30pg3^{*}$ & SE-CSLAM & 111 & - & 450 & {\color{blue}{-}} & {\color{blue}{-}} \\
   $\tt Arctic$ & SE & 111 & 28 & 450 & {\color{blue}{15947.41}} & {\color{blue}{16675.45}} \\
   $\tt Arctic-GrIS$ & SE & 111 & 14 & 225 & {\color{blue}{40305.03}} & {\color{blue}{41036.67}} \\
 \hline
 \end{tabular}
  \caption{Grids and dycores used in this study. $\Delta x_{\mathrm{eq}}$ is the average equatorial grid spacing, $\Delta x_{\mathrm{fine}}$ is the grid spacing in the refined region (if applicable), and $\Delta t_{\mathrm{phys}}$ is the physics time step. FV refers to the finite-volume dycore, SE the spectral-element dycore, and SE-CSLAM the spectral-element dycore with CSLAM tracer advection.  We use the $\tt ne30pg3$ grid for two runs with different values of $\Delta t_{\mathrm{phys}}$. {\color{blue}{The last columns provide the computational costs in core hours per simulated year (CHPSY). The costs are from single month runs using 25 nodes and 50 nodes on the Cheyenne supercomputer \cite{cheyenne}.}}}
 \label{tbl:table1}
 \end{table*}

@@

\begin{figure}[t]
\begin{center}
         \includegraphics[width=130mm]{temp_contours_diffCALIPSOdiffCERESdiffRACMO.pdf}
\end{center}
\caption{1979-1998 northern hemisphere summer surface incident shortwave radiation bias (W/m2), computed as the difference (top) from RACMO2.3p2, and (middle) the CERES-EBAF dataset, and the (bottom) total cloud fraction bias relative to the CALIPSO dataset. CALIPSO and CERES differences are found by mapping the model output to the $1^{\circ}$ grid, and differences in the bottom panel are computed after mapping the RACMO2.3p2 dataset to the individual model grids. Note that the averaging period for the CALIPSO-GOCCP and CERES-EBAF panels, 2006-2017 and 2003-2020, respectively, are different from the averaging period for the model results.}
\label{fig:FSDS}
\end{figure}

\begin{figure}[t]
\begin{center}
         \includegraphics[width=130mm]{temp_contours_diffRACMO}
\end{center}
\caption{1979-1998 (top) annual precipitation and (bottom) ice/snow melt biases relative to RACMO2.3p2, evaluated on the native model grids. The precipitation biases are expressed as fractional changes, whereas the melt biases are absolute changes (mm/yr). In the bottom panel, the \citeA{RM2012GRL} basin boundaries are shown in grey for each model grid. Note that Figure~\ref{fig:basin} uses the basin boundaries for the two common ice masks, shown in the $\tt f19$ and $\tt ne30pg2$ panels, in computing the basin-scale integrals. Blue lines in the $\tt f19$ panel show the location of the two transects plotted in Figure~\ref{fig:ztrans}.}
\label{fig:prect}
\end{figure}

@@

The bottom panel of Figure~\ref{fig:prect} presents the biases in the combined ice/snow melt as map plots. These plots show that the largest melt biases are on the southeast and northwest coasts, where large coarse-grid cells overlap with the ocean. One possibility is that these problematic grid cells are situated at lower elevations than the true ice sheet surface, leading to a warm bias and too much melt. Figure~\ref{fig:ztrans} shows the representation of the ice sheet surface along two transects on the different grids, compared to the high-resolution dataset used to generate CAM topographic boundary conditions \cite{GMTED2010,LetAl2015GMD}. The two transects are shown in Figure~\ref{fig:prect}: the east-west ``K-transect'' in southwest Greenland and a transect extending from the central dome down to the  Kangerlussuaq glacier on the southeast coast. The $1^{\circ}-2^{\circ}$ grids are noticeably coarse, with only a handful of grid cells populating the transect. The $\tt f09$ grid is a bit of an exception since the grid cells narrow in the zonal direction at high latitudes, and so a larger number of grid cells fit into the east-west transects. The VR grids more accurately reproduce the steep margins of the ice sheet, capturing the characteristic parabolic shape of the GrIS margin.

@@

Figure~\ref{fig:bias} shows the mean melt bias, relative to both RACMO datasets, conditionally sampled based on grid cell ice fraction in the GrIS region. Errors are computed after mapping the melt rates to the common ice masks using different methods, described in section~\ref{sec:SMB}. The grid cell ice fractions therefore pertain to ice fractions on the low-resolution common ice masks. Also shown are the $\pm 1$~standard deviation of the biases for each bin. The figure shows that coarser grids can be characterized by a monotonic increase in melt bias as the grid cell ice fraction decreases. The VR grids have the smallest melt biases for small grid cell ice fractions (less than 50\%), the {\color{blue}{quasi-}}uniform SE grids and $\tt f19$ have the largest melt biases and the $\tt f09$ grid melt biases lie between these two cases. Figure~\ref{fig:bias} generally supports our hypothesis that the prevalence of mixed-grid cells in the ablation zone of coarse grids is responsible for their large melt bias.

@@

{\color{blue}{Another potential factor in explaining the differences across the various grids and dycores is the surface albedo. Bare ice has a lower albedo than snow, so it's reasonable to speculate whether the large melt biases in the coarse grid simulations are in part due to more frequent bare ice exposure. One logical pathway to more bare ice exposure is related to the precipitation and cloud biases. As the coarser models accumulate too much precipitation in the GrIS interior, the coastal ice margins receive too little snow, which can be more easily melted away to expose the underlying glacial ice. Figure~\ref{fig:alb} shows maps of summer albedo biases in the simulations, relative to the RACMO2.3p2 and CERES-EBAF datasets. The surface albedo is diagnosed from the incident and net shortwave radiative fluxes at the surface. The lat-lon and quasi-uniform grids all have anomalously low albedo around the coasts in the southeast and northwest regions of the GrIS, which coincides with regions of anomalously low cloud cover and precipitation. In the VR simulations, the low albedo biases are alleviated to a large extent.

In light of these albedo results, one may interpret Figure~\ref{fig:bias} differently. That there are larger melt biases for grid cells with only only partial ice cover, may have less to do with the partial ice cover and more to do with the enhanced bare ice exposure in the coarser grids. Further experimentation is required to tease out the relative contributions of these two processes.}}

\begin{figure}[t]
\begin{center}
\begin{tabular}{cccc}
         \includegraphics[width=130mm]{temp_contours_diffALB.pdf}
\end{tabular}
\end{center}
\caption{1979-1998 northern hemisphere summer albedo bias (fraction), computed as the difference (top) from RACMO2.3p2, and (bottom) CERES-EBAF dataset. RACMO2.3p2 differences are found by mapping the model output to the individual model grids, and differences in the bottom panel are computed after mapping the individual model grids to the CERES-EBAF $1^{\circ}$ grid. Note that the averaging period for CERES-EBAF panels are 2003-2020, which is different from the averaging period for the model results.}
\label{fig:alb}
\end{figure}

@@

As modeling systems move away from lat-lon grids towards quasi-uniform unstructured grids, it is worth taking stock of whether this will degrade the simulated polar climate. We have found that the $1^{\circ}$ FV model has clear advantages over the $1^{\circ}$ SE model for simulating the GrIS SMB. {\color{blue}{The SE dycore is still under active development (e.g., \ref{sec:a1}) compared to the more mature FV dycore, and future algorithmic improvements may reduce the FV-SE GrIS SMB skill gap. However, such developments are unlikely to eliminate this skill gap entirely due to the fewer number of grid cells representing high latitude structures on quasi-uniform unstructured grids.}} The simulated GrIS SMB will be adversely impacted in future CESM versions, after the FV dycore is phased out. This finding will not interrupt the ongoing transition towards unstructured grids in CESM, which is largely driven by gains in computational efficiency and grid refinement capabilities. We therefore provide the Arctic refined-meshes to the community by way of CESM2.2, providing users the option to simulate a more realistic GrIS SMB, although at a substantial computational premium relative to conventional $1^{\circ}-2^{\circ}$ grids.
